\documentclass{article}
\usepackage[utf8]{inputenc}
\input{commands}
\input{theoremstyle.tex}

\title{Week 1}
\author{\aut}
\begin{document}

\maketitle
\section{Introduction}  

What is the course about?\\

Roughly: solving linear systems of the form
$$Ax=y$$
where \func{A}{X}{Y} is linear, $y\in Y$ given, find solutions $x\in X$. $X$ and $Y$ are linear spaces.
\begin{remark}
	When $X$ and $Y$ are finite dimensional spaces, this is linear algebra.\\
	$\infty$-dim brings into play additional structure, then it's functional analysis. (completeness, compactness, metric, norm)
\end{remark}

Example: $f\in C_0^\infty (\real^n)$, solve
$$
	\underbrace{-\Delta}_A u=f\in\real^n
$$where
$$
	C_0^\infty(\real^n)=\left\{
	f:\real^n\to\real:f\text{ infinitely differentiable }, supp(f)=\{x:f(x)\neq0\} \text{ compact }
	\right\}
$$
What space?
Can take (in PDE) $X=Y=C_\infty(\real^n)$: function spaces.
Adequate choice of space to find solutions necessary!
con vary, metric/normed linear space, locally convex topological space...

In this course: linear space is (almost always): Banach space or Hilbert space.
\begin{example}[Running example, $L^P$]
	cf. MATH50006 notes $\S2.6$ .
	$(X,\mathcal{A},\mu)$:measure space.
	$$
		L^P(\mu)\equiv l^P(X,\mathcal{A},\mu)=
		\left\{
		f:X\to\mathbb{K}/\sim:\norm{f}_{L^P}<\infty
		\right\},\quad p\in[1,\infty]
	$$
	Norm of $f\in L^p$:
	$$
		\norm{f}_p\equiv\norm{f}_{L^p}=
		\left\{
		\begin{aligned}
			 & \left(\int|f|^p{\rm d}\mu\right)^{1/p} & p<
			\infty                                               \\
			 & \esssup_X |f|                          & p=\infty
		\end{aligned}\right.
	$$
\end{example}


Choices of measure space:\\
\underline{Example 0}:\\ $X=\{1,2,....,n\}$, $\mathcal{A}=2^X$, $\mu(\{k\})=1, \forall k=1,...,n$, counting measure on $X$, extended to measure by additivity.\\
Every function $f:X\to\real$ is simple
$$
	f(x)=\sum^n_{k=1}f(k)1_{\{k\}}(x)\qquad 1_A(x)=
	\left\{\begin{aligned}
		 & 1 & x\in A    \\
		 & 0 & x\notin A
	\end{aligned}\right.
$$
Thus
$$
	\norm{f}_p^p=\int|f|^p{\rm d}\mu=\sum_{k=1}^n|f(k)|^p
	\underbrace{\int1_{\{k\}}{\rm d}\mu}_{\mu(\{k\})=1}
	=\sum_{k=1}^n|f(k)|
$$
So $L^p(\{1,...,n\},\mu\cong\real^n$ endowed with norm
$$
	\norm{p}_p=\left(\sum_{k=1}^n|f(k)|^p\right)^{1/p}\quad f\in\real^n
$$
which are finite dimensional vector spaces over $\real$ (linear algebra).

\underline{Example 1}:\\
Same with $X=\natu=\{1,2,3,...\}\mathrel{\leadsto} \ell^p$, "Little-l-p",\\
i.e. $\mu(\{k\})=1\,\forall k\in\natu$, extended to measure by $\sigma$-additivity, i.e.
$$
	\mu(A)=\sum_{k\in A} \mu(\{k\})=|A|\quad A\subset \natu
$$
Now every $f:X\to \real $ of form

$$f(x)=\sum_{k=1}^\infty f(k)1_{\{k\}}(x)$$
Approximated by $f_n\mathrel:=\sum_{k=1}^n f(k)1_{\{k\}}(x)$ and use of monotone convergence theorem to get
$$\norm{f}_p^p=\sum_{k=1}^\infty |f(k)|^p$$
An element $f:X\to \real$, $f=(f(1),f(2),...)\equiv(f_1,f_2,..._\equiv(f_k)_k$ is a sequence!
$$\ell^p=
	\{
	\text{all real-valued sequences }
	f=(f_k)_k
	\text{ s.t. }
	\sum_{k=1}^\infty|f_k|^p<\infty
	\}$$
Specially, when $p=1$, the space $\ell^1$ is the set of all \underline{absolutely convergent series}!.
\begin{remark}
We can add weights to the measures in examples above, the resulting space is denoted $\ell^p(\eta)$, where we define 
$\mu(\{j\})=\eta_j, \eta_j \geq 0, \forall j\in\natu$, thus the norm is $$\norm{f}_{\ell^p(\eta)}=\sum_{j=1}^\infty|f_j|^p\eta_j
$$
\end{remark}

\underline{Example 2}\\
$X=\real^n$ for some $n\in\natu$, $\mathcal{A}=$ Borel $\sigma$-algebra, $\mu=$Lebesgue meausre, we denote the space as $L^P(\real^n)$, which has norm:  

$$\norm{f}_p=\left(\int|f|^p{\rm d}x\right)^\frac{1}{p}$$  

where we are integrating w.r.t the Lebesgue measure. \\  

More generally, $x\subset \real^n$ open/closed $\mathrel{\leadsto} L^P(X)$, e.g. $n=1$, $X=[0,1]$.



\begin{theorem}
	Let $(X,\mathcal{A},\mu)$ be {\underline {any}} measure space. Then:
	\begin{itemize}
		\item [i)] $\norm{f}_p$ defines a \underline{norm} $\forall p\in[1,\infty]$, triangular inequality (Minkowski ineq.) holds: $\norm{f+g}_p\leq\norm{f}_p+\norm{g}_p$.
		\item [ii)] Holders inequality holds: if $\frac{1}{p}+\frac{1}{q}=1$ with $p,q\in[1,\infty]$, then $\forall f\in L^p(\mu),\,\forall g\in L^q(\mu)$, then $f\cdot g\in L^1(\mu)$ and $\norm{fg}_{L^1(\mu)}\leq\norm{f}_{L^1(\mu)}\norm{g}_{L^1(\mu)}$
		\item [iii)] $L^p(\mu)$ is complete
	\end{itemize}
\end{theorem}

\begin{definition}[Banach Space]
	A normed linear space $(X,\norm{\cdot}$ which is complete w.r.t the induced metric is called \textbf{Banach space}.\\
	Explanation:
	\begin{itemize}
		\item $\norm\cdot$ is a norm. See definition 3 in notes.
		\item the induced metric is $d(x,y)=\norm{x-y}$. It is a metric (proposition 2 in the notes) % d(x,y) is the notation used in year-2 analysis 
		\item $X$ complete w.r.t. $d$: every Cauchy sequence converges.
		\item Cauchy sequence: $(f_n)_n\subset X$ s.t. $\forall \varepsilon>0,\,\exists N\,\forall m,n\geq N,\,d(f_m,f_n)<\varepsilon$
	\end{itemize}
\end{definition}

\begin{remark}
	Theorem asserts $L^P(\mu)$ is a Banach space. Apply in the case of 1 to get immediately all of theorem 2, exercise 10-12 proposition 7 example 12-14 and much more
	!

\end{remark}

\underline{Furthur examples}\\
3. $C([a,b])=\{f:[a,b]\to\real,\,\text{continuous}\}$ with $\norm{f}_\infty=\sup_{x\in[a,b]}|f(x)|$\\
4. $C^r([a,b])$, similar to above but $f$ is set to be r-times continuously differentiable. In particular $C=C[(a,b)]=C^0[(a,b)]$.  $\norm{f}_{r,\infty}=\sup_{x\in[a,b],1\leq k\leq r}|f^{(k)}(x)|$\\
5. Sobolev space, for solving PDE(not this course).

\begin{proposition}
	$(C[0,1],\norm{\cdot}_\infty)$ is complete.
\end{proposition}
\underline{General strategy to show completeness of  $(X,\norm{\cdot})$}: For a given Cauchy sequence $(f_n)\subset X$
\begin{enumerate}
\item  find candidate limit $f$
\item  show $\norm{f_n-f}\xrightarrow{n\to\infty}0$
\item  show $f\in X$
\end{enumerate}
	

\begin{proof}\nl
\underline{STEP I}  Let $(f_n)\subset C$ be Cauchy sequence, i.e. $\forall \varepsilon>0,\,\exists N,\,\forall m,n\geq N:\,\norm{f_m-f_n}_\infty<\varepsilon$.\\
But $\norm{f_n-f_m}_\infty\geq|f_n(x)-f_m(x)|,\,\forall x\in[0,1]$ so $\left(f_n(x)\right)$ is Cauchy sequence in $\real$ $\forall  x\in[0,1]$. Since $\real$ is complete, it has a limit. Call it $f(x)=\lim_{n\to\infty}f_n(x)$.\\

\underline{STEP II} 
$|f_n(x)-f_m(x)|<\varepsilon\,\forall n,m\geq N,\,\forall x\in[0,1]$, which implies
$\lim_{n\to\infty}|f_n(x)-f_m(x)|<\varepsilon$, we also have that $lim_{n\to\infty}|f_n(x)-f_m(x)|=|f(x)-f_m(x)|$ by continuity of $f$.\\

\underline{STEP III} 
To show $f\in C$, need to argue 
\[
\forall x,\,\varepsilon>0,\,\exists\delta:\,|x-y|<\delta\implies |f(x)-f(y)|<\varepsilon \tag{$C1$}\label{C1}
\]
Now we use $\varepsilon/3$ argument. Write for any $n\in\natu$ and $x,y\in[0,1]$,
\[
	|f(x)-f(y)|\leq|f(x)-f_n(x)|+|f_n(x)-f_n(y)|+|f_n(y)-f(y)|\tag{$C2$}\label{C2}
\]
First, using STEP II, pick $n$ s.t. $\norm{f_n-f}_\infty<\frac{\varepsilon}{3}$, whence
\[
	|f(x)-f_n(x)|<\frac{\varepsilon}{3},\,|f(y)-f_n(y)|<\frac{\varepsilon}{3},\,\,\forall x,y\in[0,1]\tag{$C3$}\label{C3}
\]
Then, with $n$ now fixed, using continuity of $f_n$, pick $\delta>0$ s.t.
\[
	|f_n(x)-f_n(y)|<\frac{\varepsilon}{3}\text{ whenever } |x-y|<\delta\tag{$C4$}\label{C4}
\]
Substitute \Cref{C4},\Cref{C3} into \Cref{C2} to get \Cref{C1}.

\end{proof}


\begin{remark}
	\begin{itemize}
		\item \cref{C3} is well-known from analysis I-II. $\lim_{n\to\infty}\norm{f_n-f}_\infty=0$ is precisely the \underline{uniform convergence} of $(f_n)$ towards $f$. so 3 asserts that "uniform limit of a sequence of continuous functions is again continuous".
		\item $f_n(x)=x^n$ is not a Cauchy sequence in $(c,\norm{\cdot}\infty$, yet $f_n(x)\xrightarrow{n\to\infty} f(x)=1_{\{1\}}(x)$.
		\item if instead consider $f_n(x)=x^n$ to be elements of $L^1([0,1])$:
		      $$\int_0^1f_n{\rm d}x=\left.\frac{x^{n+1}}{n+1}\right|^1_0=\frac{1}{n+1}<\infty$$
		      Then $(f_n)\subset L^1([0,1])$ a Cauchy sequence(Ex.), and converges by  completeness of $L^p(\mu)$. The limit in $L^1([0,1])$ is $f=0$:
		      $$\norm{f_n}_{L^1([0,1])}=\norm{f_n-0}_{L^1([0,1])}=(n+1)^{-1}\xrightarrow{n\to\infty}0$$
	\end{itemize}

\end{remark}
\end{document}

