\documentclass{article}

\input{commands.tex}

\title{Functional Analysis Chapters 4 and 5}

\date{\today}

\begin{document}

\maketitle

\section{Duality}  

Recall that the space of all bounded linear operators is defined as

$$
\mathcal{L}(X,Y) = \{A: X \to Y, A \textrm{ bounded, linear}\}
$$ 

$\mathcal{L}(X,Y)$ is Banach if $Y$ is Banach and it has norm  

$$
\norm{A} = \norm{A}_{\mathcal{L}(X,Y)} = \sup_{\norm{x}_X \leq 1} \norm{Ax}
$$  

An Important special case is  

$$
X^* \overset{\textrm{def}}{=} \mathcal{L}(X,\real)
$$  

which is the \textit{dual space} of $X$.  

\begin{definition}
    (Dual Spaces) The space of all continuous linear operators $\mathcal{L}(X, \real)$ is called the \textbf{dual space} of $X$ and is denoted as  $X^*$
\end{definition}  

\begin{remark}
    $X^*$ is always Banach (eventhough $X$ may not be). We often abbreviate $\norm{\cdot}_* = \norm{\cdot}_{X^*}$. The elements of $X^*$ are called (bounded, linear) \textbf{functionals}.
\end{remark}

Dual spaces play a central role in functional analysis. They are easiest to grasp in the following contexts.  

\subsection{Duality in Hilbert Spaces}  
In this section, let $(H, {\langle \cdot, \cdot \rangle}_H )$ be a Hilbert space over $\real$. For $y\in H$, we define the map  

\begin{equation*}
    \Lambda_y: X \to \real, \qquad x \mapsto {\langle y, x \rangle}_H
\end{equation*}  

We note that this is an injective map from $H$ to its dual $H^*$ and we will show that this is in fact a bijective isometry.  

\begin{lemma}
    (Mapping to dual space)
    \begin{enumerate}[i)]
        \item $\Lambda_y \in H^*$
        \item The map $\Lambda: H \to H^*$ is a linear isometry with $||\Lambda_y||_{*}=||y||$
    \end{enumerate}
\end{lemma}  

\begin{proof}
    \begin{enumerate}[i)]
        \item We need to check the linearity and boundedness of the operator $\Lambda_y^*$. The former follows from that of the inner product and the latter is proven by applying Cauchy-Schwarz  
        \begin{equation*}
            ||\Lambda_y||_* = \sup_{x\in H, ||x|| \leq 1} |\inne{y}{x}_H| \leq \norm{y}_H
        \end{equation*}  
        which implies $\Lambda_y \in H^*$
        
        \item Choose $x = \frac{y}{\norm{y}_H}$ to attain the equality in the equation above , whence we have $\norm{\Lambda_y}_*=\norm{y}$.
    \end{enumerate}
\end{proof}

\begin{theorem}
    (Riesz Representation)  
    For every $\ell \in H^*$, there is a unique $y \in H$, such that $\ell=\Lambda_y$
\end{theorem}  

\begin{proof}
    We show the existence and uniqueness of such a linear operator.  
    \begin{itemize}
        \item \textbf{(Existence)} If $\ell(x)\equiv 0$, then take $y=0$. Otherwise, assume $\norm{\ell}_*=1$ (as we can replace $\ell(\cdot)$ by $\frac{\ell(\cdot)}{\norm{\ell}_*}$).  
        By the definition of $\norm{\cdot}_*$, there is a sequence of $(y_n)_{n\in \natu} \subset H$ with  
        \begin{equation*}
            |\ell(y_n)| \to \norm{\ell}_*, \qquad \norm{y_n}=1, \forall n \in \natu
        \end{equation*}
        We will show that the limit of this sequence is the desired $y$.  
        
        \textbf{Claim 1:} The sequence $(y_n)_{n \in \natu}$ is Cauchy   
        
        Apply the parallelogram identity to $x = \frac{y_n}{2}$ and $y = \frac{y_m}{2}$, so we have  
        \begin{equation*}
            \forall n,m \geq 1, \qquad \norm{\frac{y_n-y_m}{2}}^2=1-\norm{\frac{y_n+y_m}{2}}^2
        \end{equation*}  
        Using linearity and boundedeness of $\ell$,   
        \begin{equation*}
            \frac{1}{2} |l(y_n)+l(y_m)| = |l(\frac{y_n+y_m}{2})| \leq \norm{l}_* \norm{\frac{y_n+y_m}{2}}
        \end{equation*}  
        The LHS of the equation above converges to $1$ by assumption on $(y_n)_{n \in \natu}$, which implies and $(y_n)_{n \in \natu}$ is Cauchy.  
        Since $H$ is complete, there is a unique $y$, such that $y_n \to y$  
        
        \textbf{Claim 2:} $\ell=\Lambda_y$  
        
        Since $\textrm{span}\{y\}$ is closed, we can consider the orthogonal decomposition $H = \textrm{span}\{y\} \oplus (\textrm{span}\{y\})^{\perp}$.  
        It suffices to show:  
        \begin{enumerate}[(1)]
            \item $\ell(y)=\Lambda_y(y), \forall y \in \textrm{span}\{y\}$
            \item $\ell(x)=\Lambda_y(x), \forall x \in (\textrm{span}\{y\})^{\perp}$
        \end{enumerate}  
        To show (1), assume wlog $\norm{y}=1$, we note by continuity of $\ell$  
        \begin{equation*}
            \ell(y) = \lim_{n \to \infty} |\ell(y_n)| = \norm{\ell}_*=1
        \end{equation*}  
        and 
        \begin{equation*}
            \norm{y}^2_H = \inne{y}{y}_H =\Lambda_y(y)=1
        \end{equation*}  
        So $\Lambda_y(y)=\ell(y)$.  
        
        To show (2), we need to argue that 
        \begin{equation*}
            \ell(x)=0, \qquad \forall x \in (\textrm{span}\{y\})^{\perp}
        \end{equation*}  
        Now take $y_a = \frac{y+ax}{\sqrt{1+a^2}}$ and $\norm{y_a}=1$, where $y \in \textrm{span}\{y\}, a\in \real$ and $x \in (\textrm{span}\{y\})^{\perp}$. By definition of the norm $\norm{\cdot}_*$ and (1),  
        \begin{equation*}
            \ell(y_a)  \leq |\ell(y_a)| \leq 1 = \ell(y)
        \end{equation*}  
        So $\ell(y_a)$ has a global maximum at $a=0$ ($y_0=y$). Therefore,  
        \begin{equation*}
            0 = \frac{d}{da} \ell(y_a) \Bigr|_{a=0} = \frac{d}{da} \frac{1}{\sqrt{1+a^2}} (\ell(y)+a\ell(x))=\ell(x)
        \end{equation*}  
        So $\ell(x)=\Lambda_y(x), \forall x \in (\textrm{span}\{y\})^{\perp}$.
        
        \item \textbf{(Uniqueness)}  
        If $\ell=\Lambda_y=\Lambda_z$ for some $y,z \in H$, then  
        $$
        \forall x \in H \qquad \Lambda_y(x) = \inne{y}{x}_H = \inne{z}{x}_H = \Lambda_z(x)
        $$  
        Pick $x=y-z$, then $\inne{y}{z-x}_H = \inne{z}{z-x}_H \implies \norm{y-z}_H^2=0$. Hence $y=z$.
    \end{itemize}
\end{proof}

\begin{corollary}
    All Hilbert spaces $H$ are isomorphic to their duals $H^*$.
\end{corollary}  

\begin{unexaminable}
\begin{example}
    Some of the examples are:  
    \begin{itemize}
        \item $(l^2)^* \cong l^2$
        \item $(L^2(\mu))^* \cong L^2(\mu)$
    \end{itemize}
\end{example}
\end{unexaminable}

\subsection{Duality in Banach Spaces}  

\begin{theorem}
    (Conjugates are duals) For all $p \in (1, \infty)$, $(\ell^p)^* \cong \ell^q$, where $\frac{1}{p}+\frac{1}{q}=1$
\end{theorem}  

\begin{proof}
    For $y \in \ell^q$ define  
    \begin{align*}
        \Lambda_y: \ell^p &\to \real \\
        x &\mapsto \Lambda_y(x) = \sum_{n \in \natu} y_n x_n
    \end{align*}
\begin{lemma} (cf. \todo{what lemma})
\label{isometry in lp dual}
\begin{enumerate}[i)]
    \item $\Lambda_y \in (\ell^p)^*$
    \item $\Lambda: \ell^q \to (\ell^p)^*, y \mapsto \Lambda_y$ is a linear isometry
\end{enumerate}  
\end{lemma}
\begin{enumerate}[i)]
    \item By Hölder's inequality,  
    	\begin{equation*}
		\left|\Lambda_y(x)\right|
		\leq\sum_{n=1}^{\infty}|y_n x_n|
		\leq\norm{y}_q\norm{x}_p
	\end{equation*}
	in particular $\Lambda_y$ is well-defined (i.e. maps to $\real$). The inequality implies $\norm{\Lambda_y}_* \leq \norm{y}_q$.   
	In fact, one has equality. Let $x=(x_n)_n$ with  
	$$x_n = \textrm{sign}(y_n) |y_n|^{q-1}$$
	where the sign function is  
	\begin{equation*}
	    \textrm{sign}(t) = \begin{cases}
	    1 & t \geq 0 \\
	    -1 & t<0
	\end{cases}
	\end{equation*}
	with $\textrm{sign}(t)t=|t|, \forall t \in \real$.  
	Then $x \in \ell^p$ as $|x_n|^p=|y_n|^{p(q-1)}=|y_n|^q$ so  
	$$
	\norm{x}_p = \left(\sum_{n \in \natu} |y_n|^q\right)^{\frac{1}{p}} = \norm{y}_q^{\frac{q}{p}} = \norm{y}_q^{q-1}. 
	$$
	and
	$$
	|\Lambda_y(x)| = |\sum_{n \in \natu} x_ny_n| = \sum_{n \in \natu} |y_n|^q = \norm{y}^q_q =\norm{y}_q\norm{x}_p 
	$$
	which implies  
	$$
        \norm{\Lambda_y}_* \geq \frac{|\Lambda_y(x)|}{\norm{x}_p} = \frac{\Lambda_y(x)}{\norm{x}_p} = \norm{y}_q
	$$
        from which ii) follows.
\end{enumerate}
To complete the proof of the theorem, we have to show the following analogue of Riesz representation theorem (which applies only when $p=q=2$).  

\begin{lemma}
    $\Lambda: \ell^q \to (\ell^p)^*$ given by $y \mapsto \Lambda_y$ is surjective (onto).
\end{lemma}  
Let $e_n = (0, \ldots, 0, 1, 0, \ldots) \in \ell^p$ and define $y_n = \ell(e_n)$ for $\ell \in (\ell^p)^*$

\textbf{Claim:} 
\begin{enumerate}[i)]
    \item $y = (y_n) \in \ell^q$
    \item $\ell = \Lambda_y$ for some $y$
\end{enumerate}  

Consider the "truncated y": $y^{(n)}=(y_1, \ldots, y_n, 0, \ldots) = \sum_{i=1}^n y_i e_i \in \ell^q$ and 
$$x^{(n)}=\sum_{i=1}^{n}|y_{i}|^{q-1} \textrm{sign}(y_i) e_i \in \ell^p$$ 

Then as before: $\norm{x^{(n)}}_p = \norm{y^{(n)}}_q^{q-1}$  
with 
$$
\ell(x^{(n)})=\sum_{i=1}^{n}|y_{i}|^{q-1} \textrm{sign}(y_i) \ell(e_i) = \sum_{i=1}^n |y_i|^q=\norm{y^{(n)}}^q_q  
$$
where $y_i = \ell(e_i)$. Hence
$$
 \left(\sum_{i=1}^n |y_i|^q \right)^{\frac{1}{q}} = \norm{y^{(n)}}_q = \frac{\ell(x^{(n)})}{\norm{y^{(n)}}_q^{q-1}} = \frac{\ell(x^{(n)})}{\norm{x^{(n)}}_p} \leq \norm{\ell}_* < \infty
$$
and letting $n \to \infty$, Claim (i) follows.  

For ii), let $x \in \ell^p$ and $\varepsilon > 0$. Since $e_n$'s form a Schauder basis, we know that  

$$\norm{x^{(n)}-x}_p \overset{n \to \infty}{\rightarrow} 0$$  

(since by definition of Schauder basis, we have unique rep. of $x = \sum_{n \in \natu} x_n e_n$)

By choosing $n$ large, we can ensure that  

$$\norm{l(x)-l(x^{(n)})}<\varepsilon/2, \qquad |\Lambda_y(x)-\Lambda_y(x^{(n)})| < \frac{\varepsilon}{2}$$  

using continuity of $\ell(\cdot)$ and $\Lambda_y(\cdot)$, where the latter is a consequence of \Cref{isometry in lp dual}.  

But writing  

\begin{equation}
|\ell(x)-\Lambda_y(x)| \leq |\ell(x)- \ell(x^{(n)})| + |\ell(x^{(n)}) - \Lambda_y(x^{(n)})|+|\Lambda_y(x^{(n)})-\Lambda_y(x)|
\end{equation}  
and observing that $\ell(x^{(n)})=\sum_{i=1}^n x_i \ell(e_i) = \sum_{i=1}^n x_i y_i = \Lambda_y(x^{(n)})$, it follows that  

$$
|\ell(x)-\Lambda_y(x)| \leq \varepsilon \quad \rm{and} \quad \ell=\Lambda_y
$$  
by letting $\varepsilon \downarrow 0$
\end{proof}

\begin{remark}
    \begin{enumerate}[1)]
        \item The proof extends to $p=1$, so $(\ell^1)^* = \ell^{\infty}$. In fact, for $(X, \mathcal{A}, \mu)$, one has  
        $$L^p(\mu)^* \cong L^q(\mu) \qquad \forall p \in [1, \infty)$$
        \item For $p=\infty$, one can still define  
        $$\Lambda: \ell^1 \to (\ell^{\infty})^* \qquad y \mapsto \Lambda_y$$
        as before and check that $\Lambda$ is a linear isometry between Banach spaces.
        
        However, it is \textbf{not} surjective.  
        
        To see this, consider 
        $$c_0 = \{(x_n): \lim_{n \to \infty} = 0\} \subset \ell^{\infty}$$
    \end{enumerate}

\textbf{Claim:} \((c_0)^* \cong \ell^1\ \)
\begin{proof}
    (Sketch) We show the following statements are true:
    \begin{enumerate}[1)]
        \item \(\Lambda_y: c_0 \to \real, x \mapsto \Lambda_y(x)\) is well-defined and bounded for any 
        $y \in \ell^1$.  
        \[|\Lambda_y(x)| \leq \norm{x}_{\infty} \norm{y}_1\]

        \item The map \( \ell^1 \to c_0^*, y \mapsto \Lambda_y\) is a linear isometry. To check 
        \(\norm{\Lambda_y}_* \geq \norm{y}_1\), use \(x = \sum_{i=1}^n \rm{sign}(y_i)e_i\), as before, where
        \(n \geq 1\).  

        Clearly, \(x \in c_0\) and  
        \[\Lambda_y(x) = \sum_{i=1}^n \rm{sign}(y_i) \Lambda_y(e_i)=\sum_{i=1}^n |y_i|\]
        so for \(y \neq 0\)  
        \[\norm{\Lambda_y}_* \geq \frac{|\Lambda_y(x)|}{\norm{x}_{\infty}}=|\Lambda_y(x)|=\Lambda_y(x)=\sum_{i=1}^n |y_i|\]

        and letting \(n \to \infty\) gives the result.  

        \item \(y \mapsto \Lambda_y\) is onto: similar as before (exercise),
    \end{enumerate}
\end{proof}
\end{remark}  

\begin{definition}
    (Dual Operators) 
    \((X, \norm{\cdot}_X)\), \((Y, \norm{\cdot}_Y)\) are normed spaces over \(\real\), and \(A: X \to Y\) is a bounded linear operator.  
    Then the dual operator 
    \[A^*: Y^* \to X^*\]
    is defined by  
    \[A^* y^* \overset{\rm{def}}{=} y^* \circ A \qquad \forall y^* \in Y^* \]
    where \(A^* y^* : X \to \real\)
\end{definition}  

A note on the notation: For \(\ell \in X^*\), instead of \(\ell(x)\), we write \(\inne{\ell}{x}\). Then the above is equivalent to
\[\inne{A^* y^*}{x} = \inne{y^*}{A x} \qquad \forall x \in X, y^* \in Y^*\]  

Later we will show that if \(A\) is a bounded linear operator, then \(A^*\) is also bounded and \(\norm{A^*}=\norm{A}\)
using Hahn-Banach; if \(X, Y\) are Hilbert spaces, then we can use the Riesz representation theorem instead.

\begin{example}
    \begin{enumerate}[1)]
        \item \(A \in \real^{m \times n}\) a real matrix, which induces linear map \(L_A: \real^n \to \real^m, L_Ax=Ax\).
              If \(A^T\) is the transpose of \(A\) and \(i_k: \real^k \to (\real^k)^* \) is the canonical isomorphisim, then
              \[(L_A)^* \circ i_m = i_n \circ L_{A^T}: \qquad \real^m \to (\real^n)^*\]
        \item More generally, if \(H\) is a Hilbert space, \(A: H \to H\) is a bounded linear operator with \(\Lambda: H\to H^*\) as the
                canonical isomorphism, the operator
                \[\tilde{A^*} \overset{\rm{def}}{=} \Lambda^{-1} \circ A^* \Lambda: H \to H\]
            is called the adjoint of \(A\) (and one writes \(A^*\) for \(\tilde{A^*}\) with abuse of notation). Thus,
            \[\inne{\tilde{A^*}y}{x} = \inne{y}{Ax}, \qquad x,y \in H\]
            where \(\inne{\cdot}{\cdot}\) is the inner product on \(H\). If \(A=\tilde{A^*}\), then \(A\) is \textbf{self-adjoint}.
    \end{enumerate}
\end{example}  

\end{document}