\documentclass{article}
\usepackage[utf8]{inputenc}
\input{commands}
\input{theoremstyle.tex}

\title{Week 5 Hahn-Banach and Consequences}

\begin{document}
\maketitle

\section{Hahn-Banach Theorem} 

\begin{definition}[Sublinear functional]\nl
\label{sublinear map}
	Let $X$ be a vector space. \func{p}{X}{\real} is called \textbf{sublinear} if the following holds
	\begin{enumerate}[i)]
		\item $p( \alpha x)= \alpha p(x),$ $\forall x\in X$ and $\alpha\geq0$
		\item $p(x+y) \leq p(x)+p(y),\,\,\forall x,y\in X$
	\end{enumerate}
\end{definition}

\begin{example}
    Any linear functional is also sublinear. Also, $p(x)=\norm{x}$ on $X$ is sublinear.
\end{example}  

\begin{theorem}[Hahn-Banach]\nl
\label{Hahn-Banach}
	Let $M\subset X$ be a linear subspace, \func{p}{X}{\real} is sublinear, and \func{f}{M}{\real} is linear with 
	\begin{equation*}
	    f(x)\leq p(x) \qquad \forall x\in M
     \tag{$*$}
	\end{equation*}

	Then, there exists a linear map \func{F}{X}{\real} with $F|_M=f$ and
	
	$$F(x)\leq p(x) \qquad \forall x\in X$$
\end{theorem}

\begin{remark}[Geometric Intuitions]
    Take $x=\real^n$, $0 \in M \subset X$ an open convex subset, if $x_1 \notin <M$, then one can find a $f: X \to \real$ linear "separating" $x_1$ from $M$:

    \textit{i.e.} $$
    \begin{cases}
        f(x) < a & , x\in M \\
        f(x) \geq a & x \notin M
    \end{cases}
    $$
    for some $a\neq 0$. One can assume $a=1$ by replacing $f$ by $\frac{f}{a}$.  

    To find $f$, we introduce  
    $$
    p(x) \defeq \inf \{r > 0: \frac{x}{r} \in M\}
    $$
    \begin{unexaminable}
        Minkowski functional, cf. PS6
    \end{unexaminable}
    (e.g. $p(x)=\norm{x}$ if $M=\{x: \norm{x}<1\}$ in a normed space $(X, \norm{\cdot})$).

    One can check (using continuity) that $p$ is sublinear and  
    $$
    p(x) < 1 \iff x \in M
    $$
    
\end{remark}

\begin{proof}
    If $M=X$, take $F=f$. Else. choose $x_1 \notin M$ and set  
    $$
    M_1 = \{x+tx_1: x\in M, t\in \real\}
    $$
    ($M\subset X$ linear). Extend $f$ to $M_1$, such that ($*$) holds on $M_1$, then 'repeat'  

    \underline{\textbf{Step (I)}}\nl
    
    For all $x,y \in M$: 
    $$
    f(x)+f(y) \overset{\text{f \ linear}}{=} f(x+y) \overset{(*)}{\leq} p(x+y) \leq p(x-x_1)+p(x_1+y)
    $$
    hence 
    \begin{equation*}
        \forall x,y\in M: \qquad f(x)-p(x-x_1) \leq p(x_1+y) - f(y) \tag{$**$}
    \end{equation*}
    Take supremum over $x$, with $y$ fixed, we get  
    $$
    \alpha \defeq \sup_{x\in M} (f(x)-p(x-x_1)) < \infty
    $$
    Define $f_1: M\to \real$ by  
    $$
    f_1(x+tx_1) = f(x) + tx \qquad ,x\in M, t\in \real
    $$
    \begin{lemma}
    \phantom{something}
        \begin{enumerate}[1)]
            \item $f_1$ is linear
            \item $f_1 \mid_M=f$ 
            \item $f_1(x) \leq p(x), \forall x \in M_1$, (\textit{i.e.} ($*$) for $(f_1, M_1)$ instead of $(f, M)$)
        \end{enumerate}
    \end{lemma}
    The first two properties are easy to check.  
    
    \underline{Proof of 3):}\nl
    By definition of $\alpha$, $\forall x \in M: f(x)-\alpha \leq f(x)-( f(x)-p(x-x_1)) = p(x-x_1)$. On the other hand, taking supremum over $x$ in $(**)$ yields, for all $y \in M$  
    $$
    f(y) + \alpha \leq p(y+x_1)
    $$  
    Overall, we have  
    $$
    \forall x\in M: \qquad f_1(x\pm x_1) = f(x) \pm \alpha \leq p(x \pm x_1)
    $$  
    Apply with $t^{-1}x(\in M)$ for $t > 0$ in place of $x$ and multiply both sides by $t$ to find  
    $$
    \forall x \in M, t>0: \qquad f_1(x \pm tx_1) \leq tp(t^{-1}x\pm x_1)=p(x\pm tx_1)
    $$
\end{proof}

NB: If $X$ has a countable basis $\{e_1, i\geq 1\}$ [e.g. $\ell^p, p\in [1,\infty)$], then we can take $x_1=e_1$ and proceed by induction. For the general case, we use:

\underline{\textbf{Step (II)}}\nl
\begin{proof}(continued:)
\begin{lemma}[Zorn's lemma]\nl
\label{Zorn's Lemma}
	Let $(P, \leq), P\neq \emptyset$ is a nonempty partially ordered set and every totally ordered subset has an upper bound, then $P$ has a maximal element.
\end{lemma}  
\begin{unexaminable}
    \begin{definition}
[Partial Order]\nl
	A \textbf{partial order} on set $X$, is a binary relation, written generically $\leq$, satisfying following property.
	\begin{itemize}
		\item transitivity: if $a\leq b$ and $b\leq c$ then $a\leq c$
		\item reflexivity: $a\leq a$
		\item anti-symmetry: if $a\leq b$ and $b\leq a$ then $a=b$

	\end{itemize}
	If we also have that for any $a$ and $b$, either $a\leq b$ or $b\leq a$, then we say $\leq$ is a total order.
\end{definition}

\begin{definition}
[Upper bound]\nl
	Let $X$ be a set partially ordered by $\leq$ and $Y\subset X$, we say an element $x\in X$ is an {\bf upper bound} of $Y$ if $y\leq x$ for all $y \in Y$

\end{definition}

\begin{definition}
[Maximal element]\nl
	Let $X$ be a set partially ordered by $\leq$ and $Y\subset X$. say $x\in X$ is a \textbf{maximal element} of $X$ if $x\leq m$ implies $m=x$.

\end{definition}
\end{unexaminable}
Take 
$$
P = \{(N,g): N\subset X \ \text{linear\ subspace}, g: N \to \real \ \text{linear}, \ g\mid_N=f, \ g\mid_N \leq p\}
$$  
and define  
$$
(N,g) \leq (O, h ) \overset{\text{def.}}{\iff} N \subset O, h\mid_N = g
$$
Then $(P, \leq)$ is partially ordered, $(M,f) \in P$ so $P \neq \emptyset$. Assume $(N_i, g_i)_{i \in I}$ totally ordered. Set $N = \cup_{i\in I} N_i$ and for $x\in N$,  
$$
g(x) = g_i(x) \qquad \text{if\ }x\in N_i
$$
Then $(N,g) \in P$. Indeed $N\subset X$ is linear, and $g$ is well-defined and linear with $g\leq p$ on $N$.
\begin{itemize}
    \item (Well-defined) If $x \in N_i \cap N_k$, $N_i \subset N_k$, then $g_k \mid_{N_i}=g_i$, hence $g_i(x)=g_k(x)$
    \item (Linear) If $x,y \in N$, then $x\in N_i, y\in N_k$, for some $i,k \in I$ and $N_i \subset N_k$ (or vice versa), so $x, y \in N_k$ and  
    $$
    g(x+y) = g_k(x+y) = g_k(x)+g_k(y) = g(x) + g(y)
    $$
    \item (Bounded by $p(x)$) Similarly, one can check $g \leq p$ on $N$ (exercise)
\end{itemize}  

$(N,g)$ is an upper bound for $(N_i, g_i)_{i \in I}$, since $N_i \subset N$ and $g\mid_{N_i} = g_i$ by definition  

So \Cref{Zorn's Lemma} applies and yields that $(P, \leq)$ has a maximal element $(N,g)\in P$. Set $F=g$.  
By definition of $P$, all properties required of $F$ hold and $N=X$. For, otherwise, one can apply \underline{\textbf{Step(I)}} to $(N,g)$ and find $(N_1, g_1) \in P$ with $(N,g) < (N_1, g_1)$, which violates the maximality of $(N,g)$.
\end{proof}


\begin{remark}

\end{remark}  

\subsection{Applications of Hahn-Banach (H-B)}  

Let $(X, \norm{\cdot}_X)$ be a normed vector space, we have the following corollaries.  

\begin{corollary}
\label{same norm extension}
Let $M \subset X$ be a linear subspace, $f$ a bounded linear functional on $M$. Then $\exists F \in X^*$ with 
$$F |_M = f \ \rm{and} \ \norm{F}_{X^*}=\norm{f}_{M^*}$$
\end{corollary}

\begin{proof}
Define \func{p}{X}{\real} via  

$$p(x) = \norm{x}_X \norm{f}_{M^*}$$  

Note that $p$ is sublinear and $\forall x \in M$ and,  

$$f(x) \leq |f(x)|=\norm{x}_X \frac{|f(x)|}{\hphantom{x} \norm{x}_X} \leq \norm{x}_X \norm{f}_{M^*}=p(x)$$  
Now apply Hahn-Banach to obtain $F: X \to \real$, with  
\begin{unexaminable}
$$\norm{F}_{X^*}(x) \leq \norm{x}_X \norm{f}_{M^*} \implies \norm{F}_{X^*} \leq \norm{f}_{X^*}$$
and the other direction of the inequality follows as $F |_M = f$
\end{unexaminable}
\end{proof}  

\begin{theorem}
\label{existence of functional equals norm squared}
	Let $X$ be a normed linear space. $\forall x\in X,\,\exists x^*\in \dual{X}$ s.t. $$\inne{\dual{x}}{x}\equiv\dual{x}(x)=\norm{x}^2_X=\norm{\dual{x}}^2_{\dual{X}}$$
\end{theorem}

\begin{proof}
	Let $M=\text{span}(x)$. Define $f: M \to \real$
	
	$$f(tx)=t\norm{x}^2_X \qquad \forall t\in \real$$
	
	Then $f$ is linear, and 
	
	$$\norm{f}_{\dual{M}} = \sup_{\norm{tx}_X \leq 1} |f(tx)| = \norm{x}_X$$
	
	Then we apply \Cref{same norm extension} to extend $f$ to $\dual{x}=F\in \dual{X}$, with $\norm{\dual{x}}_{\dual{X}}=\norm{f}_M=\norm{x}_X$ and $\inne{\dual{x}|_{M}}{x}=f(x)=\norm{x}_X^2$
\end{proof}

\begin{remark}
    \Cref{existence of functional equals norm squared} gives dual characterisation of the norm later.  
    \begin{unexaminable}
        When the space is a Hilbert space, this theorem becomes Riesz representation theorem. (without changing notation! That's why bracket is a good notation here) In short, this theorem says that you can always find a linear functional such that for its value for a chosen element is precisely the norm of this element.
    \end{unexaminable}
\end{remark}  

Using Hahn-Banach, one can "separate" all sorts of things. Two examples:  

\paragraph{1) Separating points}  

\begin{proposition}
$\forall x,y \in X$, $x\neq y$, there exists $\ell \in \dual{X}$, such that $\ell(x) \neq \ell(y)$
\end{proposition}  

\begin{proof}
    Choose $\ell \in \dual{X}$ according to \Cref{same norm extension} with $y-x$ in place of $x$.  
    Then  
    $$\ell (x-y) = \ell(x) - \ell(y) = \norm{y-x}^2_X >0$$ 
    Thus, $\ell(x) \neq \ell(y)$.
\end{proof}

\paragraph{2) Separating points from closed subspaces (Urysohn-type result)}

\begin{theorem}
\label{analog of Urysohn}
$M \subset X$ linear, closed. Assume $x_0 \in X\setminus M$, such that  
$$d = \rm{dist} (x_0, M) = \inf_{x \in M} \norm{x_0-x}_M >0$$
Then $\exists \ell \in \dual{X}$ with $\ell|_M=0$ and  
$$\norm{\ell}_{\dual{X}}=1, \  \ell(x_0)=d$$
\end{theorem}

\begin{proof}
(Sketch)
Let $M_0=\{x+t x_0: x\in M\}$. Define a linear functional \func{f}{M_0}{\real}, 
$$f(x+tx_0)=td$$

Then $f\mid_M=0$ and $f(x_0)=d$. Check $\norm{f}_{M_0^*}=1$ (exercise).  

Now apply \Cref{same norm extension} to obtain extension $\ell \overset{\text{def.}}{=}F \in X^*$ with $\norm{\ell}_{X^*}=1$.
\end{proof}

The previous Theorem \Cref{analog of Urysohn} has a lot of mileage. For instance, one get  

\paragraph{3) Alternative for \Cref{existence of functional equals norm squared}}  
\begin{proposition}
    Let $X$ be a normed linear space. $\forall x\in X,\,\exists x^*\in \dual{X}$ s.t. $$\inne{\dual{x}}{x}\equiv\dual{x}(x)=\norm{x}_X=\norm{\dual{x}}_{\dual{X}}$$
\end{proposition}
\begin{proof}
    (Sketch) Apply \Cref{analog of Urysohn} with $M = \{0\}$, $x_0=\frac{x}{\norm{x}_X}$ to recover \Cref{existence of functional equals norm squared} with $x^*(x) \defeq \norm{x}_X$.
\end{proof}

\paragraph{4) $X^*$ separable $\implies$ $X$ separable}  
\begin{proof}
    See notes Page 82, which uses \Cref{analog of Urysohn}.
\end{proof}

In particular, this gives  

\begin{corollary}
    $(\ell^{\infty})^* \ncong \ell^1$
\end{corollary}
\begin{proof}
    We know $\ell^1$ is separable (use the set $\{\sum_{i=1}^n x_i e_i: x_i \in \qq, n \in \natu\}$), so if $(\ell^{\infty})^* \cong \ell^1$, then $(\ell^{\infty})^*$ is separable and so is $\ell^{\infty}$ by the theorem above. But $\ell^{\infty}$ is not separable.
\end{proof} 

From \Cref{existence of functional equals norm squared}, one gets a dual characterization of the norm:  

\begin{corollary}
\phantom{something}
\label{duality of norm}
\begin{enumerate}[i)]
    \item $\forall x \in X$: $\norm{x}_X = \underset{{\norm{x^*}_{\dual{X}} \leq 1}}{\sup} |\inne{\dual{x}}{x}|$
    \item $\forall \dual{x} \in \dual{X}$: $\norm{\dual{x}}_{\dual{X}} = \underset{{\norm{x}_X \leq 1}}{\sup} |\inne{\dual{x}}{x}|$
\end{enumerate}
The supremum in i) is always achieved.  
\end{corollary}


\begin{proof}
For $x=0$, the RHS of i) is $0$ by linearity. Let $x \neq 0$, we show two directions of inequality.  
\begin{itemize}
    \item "$\geq$": By homogeneity, we can assume $\norm{x}_X=1$. If $\dual{x} \in \dual{X}$ satisfies $\norm{x^*}_{\dual{X}} \leq 1$, then  
    $$|\inne{\dual{x}}{x}| \leq \norm{x^*}_{\dual{X}} \norm{x}_X \leq \norm{x}_X$$
    \item "$\leq$": By \Cref{existence of functional equals norm squared}, $\exists \dual{x} \in \dual{X}$, such that $|\inne{\dual{x}}{x}|=\norm{x}_X^2=1$. So the supremum is achieved.
\end{itemize}   

For ii), note that this is the definition of operator norm.   
\end{proof}

Another consequence of \Cref{existence of functional equals norm squared} is (cf. notes Theorem 23 on Page 67 for the special case $X=Y$ Hilbert, in which case the use of Hahn-Banach can be substituted by the Riesz representation theorem.)

\begin{theorem}
Let $X,Y$ be normed linear spaces and $A \in \blf{X}{Y}$. The dual operator \func{A^*}{\dual{Y}}{\dual{X}} is bounded and $\norm{\dual{A}}_{\blf{\dual{Y}}{\dual{X}}}=\norm{A}_{\blf{X}{Y}}$
\end{theorem}  

\begin{proof}
\begin{align*}
    \norm{\dual{A}} &\stackrel{\textrm{def \ of} \ \norm{\cdot}}{=} \sup_{\norm{\dual{y}}_{\dual{Y}}} \norm{\dual{A} \dual{y}}_{\dual{X}} \\
    &\stackrel{\textrm{def \ of} \  \norm{\cdot}_{\dual{X}}}{=} \sup_{\norm{\dual{y}}_{\dual{Y}}} \sup_{\norm{x}_X=1} |\inne{\dual{A}\dual{y}}{x}| \\
    &\stackrel{\textrm{def \ of} \ \dual{A}}{=} \sup_{\norm{x}_X=1}  \sup_{\norm{\dual{y}}_{\dual{Y}}}  |\inne{\dual{y}}{Ax}| \\
    &\stackrel{\phantom{\textrm{def \ of} \ \dual{A}}}{=} \sup_{\norm{x}_X=1} \norm{Ax}_Y
\end{align*}
where in the last step we used the "$\leq$" direction in the proof of \Cref{duality of norm} holds and the supremum over $\dual{y}$ is attained by \Cref{existence of functional equals norm squared}.  
\end{proof}

\end{document}