\documentclass{article}
\usepackage[utf8]{inputenc}
\input{commands}
\input{theoremstyle.tex}

\title{Week 8}

\begin{document}
\maketitle
% This file is made purely according to Written note
\section{Open mapping theorem}
\begin{definition}[Open Ball]\nl
An open ball in normed linear space $X$ with radius $r>0$ centered at $x\in X$ is
$$
B_X(x,r)=\{y\in X:\norm{y-x}_X<r\}
$$
Also, when $x=0$ we write 
$$
B_X(0,r)\equiv B_x(r)
$$
\end{definition}

\begin{definition}[Open map]\nl
	Let $X$, $Y$ be linear spaces. \func{A}{X}{Y} is {\underline{open}} if $A(U)\subset Y $ is open.
\end{definition}
\begin{remark}
\phantom{something}
\begin{itemize}
    \item $A$ being continuous means $A^{-1}(V)\subset{X}$ open $\forall V\subset Y$ open.
    \item $A$ being continuous need not be open. e.g. $Ax \defeq 0\in Y$
\end{itemize}
\end{remark}

\begin{theorem}[Open Mapping Theorem]\nl
\label{open mapping theorem}
	Let $X,Y$ be Banach, $A\subset\mathcal{L}(X,Y)$. Then:
	\begin{itemize}
	    \item[i)] if $A$ is surjective, $A$ is open.
	    \item[ii)] if $A$ is bijective, then $A^{-1}\in \mathcal{L}(X,Y)$. (Inverse operator theorem)
	\end{itemize}
\end{theorem}

\begin{remark}\nl
ii) is important in application. If $A\in\blf{X}{Y}$ is bijective then \func{A^{-1}}{X}{Y} linear is easy (why?). The point is $A^{-1}$ is also bounded, or equivalently continuous.
\end{remark}

The main step of the proof is the following:
\begin{lemma}[$A$ as in i)]\nl
\label{open ball is in closure}
$\exists r>0$ s.t. $B_Y(r)\subset \overline{A(B_X(1))}$
\end{lemma}
\begin{proof}
Since $A$ is surjective, 
$$
Y=\bigcup_{k=1}^\infty A(B_X(k))
$$
Since $Y$ is complete, by  Baire Category theorem, $\exists k_0$ s.t. 
$$ \interior (\overline{A(B_X(k_0))})\neq\emptyset$$
So by surjectivity of $A$, one can find $y_0=Ax_0\in Y$, $r_0>0$ s.t. 
$$ \underbrace{B_Y(y_0,r_0)}_{=Ax_0+B_Y(r_0)}\subset \overline{A(B_X(k_0))}$$
By linearity of $A$,
\begin{equation}\nonumber
    \begin{split}
        B_Y(r_0)&\subset\overline{A(B_X(k_0))}-Ax_0 \\ &=\overline{A(B_X(k_0)-x_0)}\\
        &\subset \overline{A(B_X(k_0+M))}   \\
        &=(k_0+M)\overline{A(B_X(1))}\\
    \end{split}
\end{equation}
Where $M\defeq \norm{x_0}_X$. So pick $r=\frac{r_0}{k_0+M}$.
\end{proof}


Proof of \Cref{open mapping theorem}:  

\begin{proof}
i) Pick $r$ as in \cref{open ball is in closure}.\\
\boxed{\underline{\textbf{Claim:}} \qquad B_Y(r/2)\subset A(B_X(1)))}\\
If claim holds, then for $U\subset X$ open, pick $x_0\in U$, $s>0$ small so that $B_X(x_0,s)\subset U$. Letting $y_0\defeq{=}Ax_0$, get 
$$
B_Y(y_0,\frac{rs}{2})=y_0+sB_Y(\frac{r}{2})\stackrel{claim}{\subset}Ax_0+sA(B_X(1))\stackrel{linearity}{=}A(B_X(x_0,s))\subset A(U)
$$
which proves i). To see i) $\implies$ ii), it's enough to show that $B=A^{-1}:Y\to X$ is continuous; but for any $U\subset X$ open, $B^{-1}(U)=(A^{-1})^{-1}(U)=A(U)$
which is open by i).
\end{proof}

\underline{\textbf{Proof of claim:}}

\begin{proof}
Fix $y\in B_Y(r/2)$. Need to show: $y=Ax$ for some $x\in X$ with $\norm{x}_X<1.$\\
We construct a sequence $(x_k)\subset X$ with 
$$
\sum_{k=1}^\infty \norm{x_k}_X<1\,\,\text{and}\,\,\sum_{k=1}^\infty Ax_k\stackrel{\norm{\cdot}_Y}{\longrightarrow{}}y,\,n\to\infty
$$
By completeness of $X$, $\sum_{k=1}^\infty x_k\defeq x$ exists, $x\in B_X(1)$ and by continuity of $A$,
$$Ax=\sum_{k=1}^\infty Ax_k=y$$
By \cref{open ball is in closure} above with $\Tilde{r}=\frac{r}{2}$,  

\begin{equation*}
    \forall s>0, \qquad B_Y(s\Tilde{r})\subset \overline{A(B_X(s/2)}
    \tag{*}
\end{equation*}
Let $s=1$. Pick $x_1\in B_X(1/2)$, such that $\norm{Ax_1-y}<\Tilde{r}/2$. 

Now set $y_1=y-Ax_1$, where $(y_1 \in B_Y(\Tilde{r}/2))$. Iterate.  

Assume that for some $k \geq 1$ have $x_1,......x_k,y_1,......y_k$ s.t.

$$
\forall 1\leq\tilde{k}\leq k:\qquad \norm{x_{\Tilde{k}}}_X < 2^{-k}\qquad y_{\tilde{k}}=y_{\tilde{k}-1}-Ax_{\tilde{k}}\in B_Y(2^{-\Tilde{k}}\Tilde{r})
$$
Then using $(*)$ with $s=2^{-(k+1)}$ find $x_{k+1}\in B_X(2^{-(k+1)})$ such that
$$
y_{k+1}\defeq y_k-Ax_{k+1}\in B_Y(2^{-(k+1)}\Tilde{r})
$$
This yields $\sum_{k=1}^{\infty}\norm{x_k}_X<1$ and 
$$
y-\sum_{k=1}^n Ax_k=y_1-\sum_{k=2}^n Ax_k=...=y_n\to0\,(n\to \infty)
$$
\end{proof}

\paragraph{\underline{\textbf{1) Equivalence of norm}}}

\begin{example}[Equivalence of Norm]\nl
Let $X=Y$, with norms $\norm{\cdot}_1$ and $\norm{\cdot}_2$ and assume $\exists C>0$ s.t.
\begin{equation}
\label{equivalence of norm example}
    \norm{x}_2\leq C\norm{x}_1,\,\forall x\in X
\end{equation}

If $X$ is complete, with respect to both $\norm{\cdot}_1$ and $\norm{\cdot}_2$ then consider $A=id:(X,\norm{\cdot}_1)\to(X,\norm{\cdot}_2)$ is open by \Cref{open mapping theorem} (indeed thm applies, as $A$ is bounded by \Cref{equivalence of norm example}). 

Since $A$ is bijective, ii) gives that $A^{-1}=id:(X,\norm{\cdot}_2)\to(X,\norm{\cdot}_1)$ is bounded, i.e.
$$
\exists C': \norm{A^{-1}}_1=\norm{x}_1\leq C'\norm{x}_2
$$
so $\norm{\cdot}_1$ and $\norm{\cdot}_2$ are actually equivalent.
\end{example}

\paragraph{\underline{\textbf{2) $Y$ needs to be complete in \Cref{open mapping theorem}}}}
\begin{example}[Completeness of $Y$]\nl
Consider $X=C(=C^0[0,1])$ with $\norm{\cdot}_1=\norm{\cdot}_\infty$, $\norm{\cdot}_2=\norm{\cdot}_{L^1}$. Then $A=id:(X,\norm{\cdot}_1)\to(X,\norm{\cdot}_2)$ is continuous:
$$
\norm{Af}_2
=\norm{f}_2
=\int_0^1|f(t)|dt
\leq\norm{f}_\infty
=\norm{f}_1
$$
but not open. Else by 1),  $\norm{\cdot}_1$ and $\norm{\cdot}_2$ would be equivalent. However consider counter example:
\begin{equation}\nonumber
f_n(x)=\left\{
\begin{split}
    &{2n^2x} &x\in[0,\frac{1}{2n}]\\
    &{-2n^2x+2n} &x\in(\frac{1}{2n},\frac{1}{n}]\\
    &0 &x\in(\frac{1}{n},1]\\
\end{split}
\right.\quad\text{satisfy}\quad\norm{f_n}_2=1,\norm{f_n}_1=n\to\infty
\end{equation}
This shows $Y$ needs to be complete in theorem.
\end{example}

\paragraph{\underline{\textbf{3) $X$ needs to be complete in \Cref{open mapping theorem}}}}
\begin{example}[Completeness of $X$]\nl
This example shows completness of $X$ is also required.
Take 
$$
X=Y=\{(x_n)\in\ell^\infty:\exists N \in \natu: \ x_n=0\,\forall n\geq N\}\subset\ell^\infty
$$
with norm $\norm{\cdot}_X=\norm{\cdot}_Y=\norm{\cdot}_\infty$. This is a linear normed space. It's not complete (Exercise: show directly $\overline{X}=c_0$). Another way:
Define \func{A}{X}{X}, 
$$
Ax=(x_1,\frac{x_2}{2},\frac{x_3}{3},\underbrace{......}_{0\,\text{eventually}})\quad \text{if} \,\,x=(x_1,x_2......)
$$
Then $A$ is linear, bijective with 
$$
A^{-1}:X\to X, \,\,\,A^{-1}x=(x_1,2x_2,3x_3\underbrace{......}_{0\,\text{eventually}})
$$
and $A$ is bounded. 
$$
\norm{Ax}_\infty=\sup_{n\geq1}\frac{|x_n|}{n}\leq\sup_{n\geq1}|x_n|=\norm{x}_\infty
$$
so $\norm{A}\leq1$. But $A^{-1}$ is unbounded. 
Pick $x^{(n)}=(\overbrace{1,1,\ldots, 1,1}^{n},0,......)$ then $\norm{x^{(n)}}_\infty=1$ but $\norm{A^{-1}x^{(n)}}=n$. Hence $A^{-1}\not\in\mathcal{L}(X)$ and $X$ cannot be complete, else by \Cref{open mapping theorem} ii), $A^{-1}$ would be bounded.
\end{example}

\section{Closed Graph Theorem}
Consider $X$, $Y$ normed spaces. Often an operator $A$ not defined on all of $A$ but on a "domain" $D(A)$. So we assume that 
$D(A)\subset X$ is a linear subspace on which $A:D(A)(\subset X)\to Y$, linear is defined.
\begin{example}[running xxample]\nl
$Y=X=C=C^0[0,1]$ with $\norm{\cdot}_X=\norm{\cdot}_\infty$ and 
$$A=\frac{d}{dt}$$
with $D(A)\stackrel{eg}{=}C^1[0,1]\subset X$ or subspaces thereof. 

Prime example of  (\underline{unbounded}) operator with dense domain $D(A)$: indeed $C^1[0,1]=C$ using e.g. Weierstrass Approximation Theorem (Polynomials are already $\norm{\cdot}_\infty$-dense in C)
\end{example}
\begin{definition}[Graph of an operator]\nl
Let $X$, $Y$ be normed space, $A:D(A)(\subset X)\to Y$ . Graph of $A$ (really  of $(A,D(A))$) is the linear (!) space 
$$
\Gamma_A=\{(x,Ax):x\in D(A)\}\subset X\times Y
$$
We endowed $X\times Y$ with the norm $\norm{(x,y)}_{X\times Y}=\norm{x}_X+\norm{y}_Y$, for all $x\in X$, $y\in Y$.
    
\end{definition}
\begin{definition}[Closed Operator]
$A$ is called \underline{closed} if $\Gamma_A$ is closed in  $(X\times Y,\norm{\cdot}_{X\times Y})$
\end{definition}

\begin{example}
Let $A\in\mathcal{L}(X,Y)$ with $D(A)=X$. Then $A$ is closed.
\end{example}
\begin{proof}
	Let $(x_k,y_k)_k\subset \Gamma_A$  with $\norm{(x_k,y_k)-(x,y)}_{X\times Y}\xrightarrow{k\to\infty}0$ for some $(x,y)\in X\times Y$\\
	Need to show: $(x,y)\in \Gamma_A$ i.e. $y=Ax$.   
 
	We know $ y_k=Ax_k$ and $\norm{x_k-x}_X\xrightarrow{k\to\infty}0$, $\norm{Ax-y}_Y\xrightarrow{k\to\infty}0$
	But $\forall k\geq 1$,
	$$\norm{y-Ax}_Y\leq\norm{y-Ax}_Y+\norm{Ax_k-ax}_Y
	\leq \norm{y-Ax}_Y+\norm{A}\norm{x_k-x}_X$$
	Thus 
	$$\lim_{k\to\infty}\norm{y-Ax}_Y\leq\lim_{k\to\infty}\norm{y-Ax}_Y+\norm{A}\norm{x_k-x}_X=0$$
\end{proof}

\begin{theorem}[Closed Graph]\nl
\label{closed graph theorem}
Let $X$, $Y$ be Banach $A:X\to Y$ linear. The following are equivalent:
\begin{itemize}
    \item [i)] $A\in\mathcal{L}(X,Y)$
    \item [ii)] $A$ is closed
\end{itemize} 
\end{theorem}
\begin{proof}
    i) $\implies$ ii): see example\\
    ii) $\implies$ i): If $X$, $Y$ complete, then so is $(X\times Y,\norm{\cdot}_{X\times Y})$ (exercise). A closed means $\Gamma_A$ is closed in $(X\times Y,\norm{\cdot}_{X\times Y})$, so $(\Gamma_A,\norm{\cdot}_{X\times Y})$ is complete. Consider:
    \begin{equation}
        \begin{aligned}
            \Pi_X:\,\,\Gamma_A&\to X \qquad\qquad& \Pi_Y:\Gamma_A&\to Y\\
        (x,Ax)&\mapsto x  & (x,Ax)&\mapsto Ax\\
        \end{aligned}    
    \end{equation}
$\Pi_X$, $\Pi_Y$ are continuous with $\norm{\Pi_X},\norm{\Pi_Y}\leq 1$, $\Pi_X$ is injective, and surjective. By \Cref{open mapping theorem} ii), ${\Pi_X}^{-1}\in\mathcal{L}{(X,\Gamma_A)}$ and so
$$
A=\Pi_Y\circ \Pi_X^{-1}\in\mathcal{L}(X,Y)
$$
\end{proof}

\begin{remark}
	ii) is simpler than i), but equivalent.\\
	i) says A is continuous, i.e. if $(x_n)\subset X$, $x\in X$
	$$\norm{x_n-x}_X\overset{n\to\infty}{\longrightarrow} 0\implies\norm{Ax_n-Ax}_Y\overset{n\to\infty}{\longrightarrow}0$$
	This contains two things to check: $(Ax_n)$ converges and limit is $Ax$.\\
	ii) says $A$ is closed, i.e.
	\begin{equation}
		\left\{
		\begin{aligned}
			&\norm{x_n-x}_X\overset{n\to\infty}{\longrightarrow}0\\
			&\norm{Ax_n-y}_Y\overset{n\to\infty}{\longrightarrow}0\\
		\end{aligned}
		\right.\implies
		Ax=y
	\end{equation}
	Which is only one condition to check.
\end{remark}

\begin{example}[running example continues]\nl
	$(D(A),\norm{\cdot}_\infty)$ with $D(A)=C^1[0,1]$ is NOT Banach, and $A:D(A)\to C$ is an example of an operator which is:\\
	\textbf{Claim: }\\
	i) closed, but\\
	ii) not continuous\\
	For ii), take $f_n(t)=t^n\in D(A)$, $Af_n=nf_{n-1}$ so $\norm{f_n}_\infty=1$, $\norm{Af_n}_\infty=n\norm{f_{n-1}}_\infty=n$. So 
	$$ \sup_{f\in D(A),\norm{f}_\infty\leq1}\norm{Af}_\infty=\infty$$
	For i), if $(f_n,f_n')\to(f,g)$ in $(D(A)\times C)$ then $\norm{f-f_n}_\infty\to 0$, $\norm{f_n'-g}_\infty\to0$ but
	$$
	\forall t\in(0,1],\,\underbrace{f_n(t)}_{\overset{n\to\infty}{\longrightarrow}f(t)}
	=\underbrace{\int_0^t f'_n(x) dx}_{\overset{DCT}{\longrightarrow}\int_0^t g(x) dx}
	+f_n(0)
	$$
	so $f'=g$ by fundamental theorem of calculus(FTC), i.e. $(f,g)=(f,f')\in\Gamma_A$. 
 \begin{unexaminable}
The second convergence uses dominated convergence theorem
 \end{unexaminable}
\end{example}


\begin{corollary}[Continuous Inverse]\nl
	$X$, $Y$ Banach, $A:(DA)\subset X\to Y$ linear, closed and bijective. Then $\exists B=A^{-1}\in\mathcal{L}(Y,X)$ with $AB=id_Y$ and $BA=id_{D(A)}$.
	Proof is left as an exercise. Hint: similar to the Closed Graph \Cref{closed graph theorem}, consider $\Pi_Y:\Gamma_A\to Y$, $B\defeq\Pi_X\circ \Pi_Y^{-1}$
	
\end{corollary}

\todo{what is ?}
\begin{example}[???]\nl
	A is surjective: for $g\in C$ define $f(t)=\int_0^t g(s) ds$. Then by FTC, $Af=g$.\\
	A is not injective: $Af=A\tilde{f}\implies f=\tilde{f}+c,c\in\real$. 
	Let $D(A)\defeq C_0^1[0,1]=\{f\in C^1[0,1]:f(0)=0\}$
	Then $A:D(A)\to C$ is bijective and has continuous inverse $B=A^{-1}$ by corollary. In fact, $Bf(t)=\int_0^tf(s)ds$ with $Bf\in D(A)$.
\end{example}



\end{document}

